### 2021-11-10 paper.Rmd
Update plots and text accordingly

---
title: 'experDesign: stratifying samples into batches with minimal bias'
output:
  md_document:
    preserve_yaml: yes
  word_document:
    keep_md: yes
  pdf_document: default
  html_document:
    df_print: paged
date: "11th November 2021"
affiliations:
- name: Centro de Investigación Biomédica en Red, Enfermedades Hepáticas y Digestivas
  index: 1
- name: Institut d'Investigacions Biomèdiques August Pi i Sunyer, IDIBAPS
  index: 2
authors:
- name: Lluís Revilla Sancho
  orcid: 0000-0001-9747-2570
  affiliation: 1, 2
- name: Juan-José Lozano
  orcid: 0000-0001-9747-2570
  affiliation: '1'
- name: Azucena Salas
  orcid: 0000-0001-9747-2570
  affiliation: '2'
tags:
- R
- batch effect
- experiment design
bibliography: paper.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	error = 0,
	fig.width = 9,
	warning = TRUE,
	collapse = TRUE
)
```

# Summary

The design of an experiment is critical to its success.
Nonetheless, even when correctly designed, the process leading up to the moment of measuring a given variable is critical.
At any one of the several steps, from sample collection to measurement of a variable, various errors and problems can affect the experimental results.
Failure to take such variability into account can render an experiment inconclusive.
*experDesign* provides tools to minimize the risk of inconclusive results by assigning samples to batches to reduce potential batch effects.

# Introduction

To design an experiment that can support conclusive results upon analysis, the source of the variation between samples must be identified.
Typically, one can control the environment in which the study or experiment is being conducted.
Sometimes, however, this is not possible.
In such cases, techniques to control variations must to be applied.
There are three methods used to decrease the uncertainty of the unwanted variation: blocking, randomization and replication [@klaus2015].

**Blocking** is a method that groups samples that are equal according to one or more variables, allowing the estimation of the differences between each batch by comparing measurements within the blocks.
**Randomization** minimizes the variation in the measurements by randomly mixing the potential confounding variables.
**Replication** increases the number of samples used in an experiment to better estimate the variation of the experiment.
In some settings these techniques can be applied together to enhance the robustness of the study.

Between the designing of an experiment and the measurement of the samples, some samples might be lost, contaminated, or degraded below the quality threshold.
In addition, experiments will occasionally need to be carried out in batches.
The later might be needed for technical reasons; for example, the device cannot measure more than a given number of samples at the same time.
Practical reasons can also be a factor; for instance, it may not be possible to obtain additional measurements in the field during the allotted time.

This divergence from the original design might cause batch effects, thereby perturbing the analysis.
There are several techniques to identify and assess batch effects when analyzing an already measured experiment [@leek2010].
It would be better to avoid such batch effects before executing an experiment.
By taking into account the differences between the original design and the state before the measurement is conducted, confounding effects can be minimized.

To prevent the batch effect from confounding the analysis after the initial design of the experiment, there are two options: randomization and replication.
Randomization, consists of shuffling the samples in order to mix different attributes, which can help reduce variations across groups.
In contrast, replication helps estimate the variation of the measurements or samples, thus increasing the precision of the estimates of the true value obtained by the analysis.
Replications consist of increasing the number of measurements with similar attributes.
When a sample is measured multiple times, this is referred to as a technical replicate.
Technical replicates help estimate the variation of the measurement method, and thus the possible batch effect [@blainey2014].

Randomization and replication can be used to prevent batch effects that might confound the analysis.
By examining how the variables are distributed across each batch, proper randomization can be ensured, thus minimizing batch effects.
This is known as randomized block experimental design or stratified random sampling experimental design.

# State of the art

There are certain tools that can minimize batch effects on the R language in multiple fields, particularly for biological research [@rcoreteam2014].
Here we briefly describe the currently available packages:

-   [*OSAT*](https://bioconductor.org/packages/OSAT/), at Bioconductor, first allocates the samples from each batch according to a variable; it then shuffles the samples from each batch in order to randomize the other variables [@yan2012].
    This algorithm relies on categorical variables and cannot use numerical variables (e.g., those that are age- or time-related) unless they are treated as categorical variables.

-   [*anticlust*](https://cran.r-project.org/package=anticlust), at CRAN, divides the samples into similar groups, ensuring similarity by enforcing heterogeneity within groups [@papenberg2020].
    Conceptually it is similar to the clustering method k-means.

-   Recently, [*Omixer*](https://bioconductor.org/packages/Omixer/), a new package, has been made available at Bioconductor [@sinke2021].
    It tests whether the random assignments made by it are homogeneous by transforming all variables to numeric values and using the Kendall's correlation when there are more than 5 samples; otherwise, it utilizes the Pearson's chi-squared test.

For completeness a description and comparison of the usage of the different software packages currently available on CRAN and Bioconductor is presented below.
First, we start with some real data obtained from a survey.
This data set has three variables of interest; Sex, Smoke and Age are a mix of categorical and numeric variables.

```{r example-data, include = FALSE}
data(survey, package = "MASS")
survey$ID <- seq_len(nrow(survey))
VoI <- c("Sex", "Smoke", "Age")
n_batch <- 3
size_subset <- 96
iterations <- 1000
```

## OSAT

OSAT provides templates for plates that hold 2, 4, 8 Illumina BeadChip chips, having 24, 48 or 96 wells.
Moreover, it works for both numeric and categorical variables but OSAT might return less rows than the input provided because they might have NA value.

```{r osat, fig.cap="Sample, distributions with OSAT.", include = FALSE}
library("ggplot2")
theme_plots <- theme_minimal() + theme(text = element_text(size = 10))
theme_set(theme_plots)
library("OSAT")
library("patchwork")
gs <- setup.sample(survey, optimal = VoI)
gc <- setup.container(plate = IlluminaBeadChip96Plate, 
                      n = n_batch, batch = 'plates')
gSetup <- create.optimized.setup(sample = gs, container = gc, 
                                 nSim = iterations)
osat_report <- get.experiment.setup(gSetup)
nrow(osat_report)
osat_age <- ggplot(osat_report) +
  geom_histogram(aes(Age, fill = as.character(plates))) +
  facet_wrap(~plates) +
  labs(fill = "Batch")
osat_sex <- ggplot(osat_report) +
  geom_bar(aes(Sex, fill = as.character(plates))) +
  facet_wrap(~plates) +
  labs(fill = "Batch")
osat_smoke <- ggplot(osat_report) +
  geom_bar(aes(Smoke, fill = as.character(plates))) +
  facet_wrap(~plates) +
  labs(fill = "Batch")
```

## anticlust

anticlust does not handle all types of variables, it only accepts numeric variables.

```{r anticlust, fig.cap="Sample, distributions with anticlust.", include = FALSE}
library("anticlust")
anticlust_index <- anticlustering(
  survey[, VoI],
  K = n_batch,
  objective = "variance",
  method = "exchange",
  repetitions = iterations
)
anticlust_index <- anticlustering(
  survey[, "Age"],
  K = n_batch,
  objective = "variance",
  method = "exchange",
  repetitions = iterations
)
report_anticlust <- data.frame(Age = survey[, "Age"], 
                               batch = as.character(anticlust_index))
library("ggplot2")
anticlust <- ggplot(report_anticlust) + 
  geom_histogram(aes(Age, fill = batch)) +
  facet_wrap(~batch) +
  labs(fill = "Batch")
```

## Omixer

There is a bug in the Omixer that prevents it from working unless specific conditions are meet.
This precluded any comparisons of Omixer with other tools using the same settings.

```{r Omixer, include = FALSE}
library("Omixer")
survey$ID <- seq_len(nrow(survey))
Omixer_index <- omixerRand(survey, sampleId = "ID", 
                           block = "block", iterNum = 10, wells = size_subset, 
                           plateNum = n_batch, randVars = VoI, techVars = NULL)
Omixer_index <- omixerRand(survey, sampleId = "ID", 
                           block = "block", iterNum = 10, wells = size_subset, 
                           mask = rep(1, nrow(survey)),
                           plateNum = n_batch, randVars = VoI, techVars = NULL)
Omixer_index <- omixerRand(survey, sampleId = "ID", 
                           block = "block", iterNum = 10, wells = size_subset, 
                           mask = rep(0, nrow(survey)),
                           plateNum = n_batch, randVars = VoI, techVars = NULL)
Omixer_index <- omixerRand(survey, sampleId = "ID", 
                           block = "block", iterNum = 10, wells = size_subset, 
                           mask = rep(1, 288),
                           plateNum = n_batch, randVars = VoI, techVars = NULL)
Omixer_index <- omixerRand(survey, sampleId = "ID", 
                           block = "block", iterNum = 10, wells = size_subset, 
                           mask = rep(0, 288),
                           plateNum = n_batch, randVars = VoI, techVars = NULL)

```
These methods results on this distributions:

```{r panel1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure A: distribution of  samples according to *OSAT*, Figure B according to *anticlust* on the variables used (1: Smoke, 2: Age, 3: Sex)."}
osat_plot <- osat_smoke / (osat_age + osat_sex)
osat_plot /anticlust + plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = list(c("A.1", "A.2", "A.3", "B.2")),
                  title = "Existing solutions",
                  subtitle = "OSAT and anticlust solutions") &
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 8, hjust = 0, vjust = 0))
```

On Figure 1, we can see that the distributions is fairly similar between these two packages.

# Statement of need

Current solutions for stratifying samples to reduce and control batch effect do not work for all cases.
They are either specialized to a particular type of data, they omit some conditions that are usually met, or they only work under a specific subset of conditions.
The new package ***experDesign*** works with all data types and does not require a spatial distribution making it suitable for all kind of experiments.
This package is intended for people needing a quick and easy solution that will provide reasonable suggestions on how to best distribute the samples for analysis.

# Description

The package ***experDesign*** provides the function `design` to arrange the samples into multiple batches such that a variable's distribution remains homogeneous within each batch.
Each batch is set to have some centrality and dispersion statistics to match as closely as possible with the original input design data.
The statistics used are the mean, the standard deviation, the median absolute deviation, variables with no value number, the entropy and the independence of the categorical variables.
With each iteration if the random distribution of the sample statistics for each batch has fewer differences vis-à-vis the original distribution than the last stored sample distribution then it replaces it as the best sample distribution.
Upon completion of the iterations the best sample distribution is returned to the user.

Users can examine the following flowchart to decide what function(s) they need to use:

![Flow chart to decide which functions are needed](flow_chart.png){width="353"}

If users want a design without replicates but the batches have some spatial distribution, we must use `design` to allocate the samples on each batch, followed by the `spatial` function to randomly distribute the samples homogeneously by position within each batch.
See the example in `inspect` and the vignette:

```{r experDesign, echo=FALSE, fig.cap="Sample, distributions with anticlust.", warning=FALSE, message=FALSE}
library("experDesign")
experDesign_index <- design(survey[, VoI], size_subset = size_subset,
                  iterations = iterations, name = "")
experDesign_index_spatial <- spatial(index = experDesign_index, 
                                     pheno = survey[, VoI], 
                                     iterations = iterations, 
                                     rows = LETTERS[1:8], 
                                     columns = 1:12)
experDesign_report <- inspect(experDesign_index, survey[, VoI])
# nrow(experDesign_report)
experDesign_age <- ggplot(experDesign_report) + 
  geom_histogram(aes(Age, fill = batch)) +
  facet_wrap(~batch) +
  labs(fill = "Batch")
experDesign_sex <- ggplot(experDesign_report) +
  geom_bar(aes(Sex, fill = batch)) +
  facet_wrap(~batch) +
  labs(fill = "Batch")
experDesign_smoke <- ggplot(experDesign_report) +
  geom_bar(aes(Smoke, fill = batch)) +
  facet_wrap(~batch) +
  labs(fill = "Batch")
(experDesign_smoke/(experDesign_age + 
                      experDesign_sex )) +
  plot_annotation(title = "experDesign") +
  plot_layout(guides = 'collect')
```

The statistics of the index can be checked for multiple statistics, as shown on the help pages of `evaluate_na`, `evaluate_entropy`, `evaluate_mad`, `evaluate_sd` and `evaluate_mean`. We can also compare our results with the original distribution via `evaluate_orig`.

```{r evaluate, include = FALSE}
evaluate_na(experDesign_index, survey[, VoI])
evaluate_entropy(experDesign_index, survey[, VoI])
evaluate_mad(experDesign_index, survey[, VoI])
evaluate_sd(experDesign_index, survey[, VoI])
evaluate_mean(experDesign_index, survey[, VoI])
# All together for each batch
ei <- evaluate_index(experDesign_index, survey[, VoI])
ei[, , "SubSet1"]
evaluate_na(experDesign_index_spatial, survey[, VoI])
evaluate_entropy(experDesign_index_spatial, survey[, VoI])
evaluate_mad(experDesign_index_spatial, survey[, VoI])
evaluate_sd(experDesign_index_spatial, survey[, VoI])
evaluate_mean(experDesign_index_spatial, survey[, VoI])
# All together for each batch
ei <- evaluate_index(experDesign_index_spatial, survey[, VoI])
ei[, , "G1"]
```

```{r original, include = FALSE}
evaluate_orig(survey[, VoI])
```

In addition to distributing the samples into batches, ***experDesign*** provides tools to add technical replicates.
In order to choose them from the available samples, the function `extreme_cases` is provided.
For easier usage, the `replicates` function designs an experiment with the desired number of replicates per batch.

```{r controls, include = FALSE}
extreme_cases(survey[, VoI], size = 5)
replicates_index <- replicates(pheno = survey[, VoI], 
                               size_subset = size_subset, controls = 5, 
                               iterations = iterations)
spatial_replicate_index <- spatial(index = replicates_index, 
                                   pheno = survey[, VoI], 
                                   rows = LETTERS[1:8], columns = 1:12)
survey$ID <- seq_len(nrow(survey))
report_replicates <- inspect(i = replicates_index, 
                             pheno = survey[, c(VoI, "ID")])
report_replicates_position <- inspect(i = spatial_replicate_index, 
                             pheno = report_replicates, 
                             index_name = "position")
# Batch and position where the samples should be:
head(report_replicates_position)
# Technical replicates:
head(report_replicates_position[duplicated(report_replicates_position$ID), ])
```

***experDesign*** also provides several small utilities to make it easier to design the experiment in batches.
For instance, a function called `sizes_batches` helps calculate the number of samples in order to distribute them across the required batches.
Furthermore, `optimum_batches` calculates the minimal number of batches required:

```{r show-batches, include = FALSE}
optimum_batches(size_data = 250, size_subset = 96)
sizes_batches(size_data = 250, size_subset = 96, batches = 3)
```

In conclusion ***experDesign*** offers a fast method for preparing a batched experiment.
It can use as many numeric and categorical variables as needed to stratify the experimental design based on batches including spatial distributions.

# Acknowledgments

We are grateful to Joe Moore for English-language assistance.

# References