### 2021-11-22 README.Rmd
Increment version number to 0.1.1

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# experDesign

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/experDesign)](https://CRAN.R-project.org/package=experDesign)
[![R build status](https://github.com/llrs/experDesign/workflows/R-CMD-check/badge.svg)](https://github.com/llrs/experDesign/actions?workflow=R-CMD-check)
[![Coverage status](https://codecov.io/gh/llrs/experDesign/branch/master/graph/badge.svg)](https://codecov.io/github/llrs/experDesign?branch=master)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![DOI](https://zenodo.org/badge/142569201.svg)](https://zenodo.org/badge/latestdoi/142569201)

<!-- badges: end -->

The goal of experDesign is to help you decide which samples go in which batch, 
reducing the potential batch bias before performing an experiment.
It provides three main functions :

* `design()`: Randomize the samples according to their variables.
* `replicates()`: Selects some samples for replicates and randomizes the samples.
* `spatial()`: Randomize the samples on a spatial grid.

## Installation

To install the latest version on [CRAN](https://CRAN.R-project.org/package=experDesign) use:

```r
install.packages("experDesign")
```
You can install the development version from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("llrs/experDesign")
```

## Example

We can use the survey dataset for the examples:

```{r show}
library("experDesign")
data(survey, package = "MASS") 
head(survey)
```

The dataset has numeric, categorical values and some `NA`'s value.

# Picking samples for each batch

Imagine that we can only work in groups of 70, and we want to randomize by Sex, 
Smoke, Age, and by writing hand.  
There are `r choose(237, 70)` combinations some of them would be have in a 
single experiment all the right handed students. We could measure all these combinations
but we can try to find an optimum value.

```{r design, fig.show='hold'}
# To reduce the variables used:
omit <- c("Wr.Hnd", "NW.Hnd", "Fold", "Pulse", "Clap", "Exer", "Height", "M.I")
(keep <- colnames(survey)[!colnames(survey) %in% omit])
head(survey[, keep])

# Looking for groups at most of 70 samples.
index <- design(pheno = survey, size_subset = 70, omit = omit)
index
```

We can transform then into a vector to append to the file or to pass to the lab mate with:

```{r batch_names}
head(batch_names(index))
```

# Previous work

The CRAN task View of [Experimental Design](https://CRAN.R-project.org/view=ExperimentalDesign) includes many packages relevant for designing an experiment before collecting data, but none of them provides how to manage them once the samples are already collected.

Two packages allow to distribute the samples on batches:

- The [OSAT](https://bioconductor.org/packages/release/bioc/html/OSAT.html) package handles categorical 
variables but not numeric data. It doesn't work with our data.

 - The [minDiff](https://github.com/m-Py/minDiff) package reported in [Stats.SE](https://stats.stackexchange.com/a/326015/105234), handles both 
numeric and categorical data. But it can only optimize for two nominal criteria.
It doesn't work for our data.

 - The [Omixer](https://bioconductor.org/packages/Omixer/) package handles both 
numeric and categorical data (converting categorical variables to numeric). But both the same way either Pearson's Chi-squared Test if there are few samples or Kendall's correlation. It does allow to protect some spots from being used.

If you are still designing the experiment and do not have collected any data [DeclareDesign](https://cran.r-project.org/package=DeclareDesign) might be relevant for you.

Question in [Bioinformatics.SE](https://bioinformatics.stackexchange.com/q/4765/48) I made before developing the package.

# Other

Please note that this project is released with a [Contributor Code of Conduct](https://www.contributor-covenant.org/version/1/0/0/code-of-conduct/).
By participating in this project you agree to abide by its terms.